\documentclass[a4paper]{scrreprt}

% Uncomment to optimize for double-sided printing.
% \KOMAoptions{twoside}

% Set binding correction manually, if known.
% \KOMAoptions{BCOR=2cm}

% Localization options
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% Quotations
\usepackage{dirtytalk}

% Enhanced verbatim sections. We're mainly interested in
% \verbatiminput though.
\usepackage{verbatim}

% PDF-compatible landscape mode.
% Makes PDF viewers show the page rotated by 90°.
\usepackage{pdflscape}

% Advanced tables
\usepackage{tabu}
\usepackage{longtable}

% Fancy tablerules
\usepackage{booktabs}

% Graphics
\usepackage{graphicx}

% Current time
\usepackage[useregional=numeric]{datetime2}

% Float barriers.
% Automatically add a FloatBarrier to each \section
\usepackage[section]{placeins}

% Custom header and footer
\usepackage{fancyhdr}

\usepackage{geometry}
\usepackage{layout}

% Math tools
\usepackage{mathtools}
% Math symbols
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amsthm}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

\pagestyle{plain}
% \fancyhf{}
% \lhead{}
% \lfoot{}
% \rfoot{}
% 
% Source code & highlighting
\usepackage{listings}

% Convenience commands
\newcommand{\mailsubject}{41100 - Cryptography - Series 2}
\newcommand{\maillink}[1]{\href{mailto:#1?subject=\mailsubject}
                               {#1}}

% Should use this command wherever the print date is mentioned.
\newcommand{\printdate}{\today}

\subject{41100 - Cryptography}
\title{Series 2}

\author{Michael Senn \maillink{michael.senn@students.unibe.ch} - 16-126-880}

\date{\printdate}

% Needs to be the last command in the preamble, for one reason or
% another. 
\usepackage{hyperref}


\begin{document}
\maketitle


\setcounter{chapter}{1}
\chapter{Series 2}

\section{Basics on libraries}

\subsection{Programs and libraries}

Let $X, Y \sim Un(\{0, 1, \cdots, n-1\})$ iid. Per its definition, the output
of $L_1.\text{RAND}$ follows $x$ respectively $Y$, and the output of $L_2.\text{RAND}$ is a constant
$0$. Hence:

\begin{itemize}
	\item $P[A_1 \diamond L_1 \Rightarrow 1] = P[X = Y] = \frac{1}{6}$
	\item $P[A_1 \diamond L_2 \Rightarrow 1] = P[0 = 0] = 1$
	\item $P[A_2 \diamond L_1 \Rightarrow 1] = P[X \geq 3] = \frac{3}{6}$
	\item $P[A_2 \diamond L_2 \Rightarrow 1] = P[0 \geq 3] = 0$

\end{itemize}

\subsection{Exchangability of libraries}

\subsubsection{Bitwise complement}

Note first that $L_{\text{left}}$ produces a uniform distribution on $\{0,
1\}^n$. As the bitwise complement is a bijection of $\{0, 1\}^n$ on itself, it
follows that $L_{\text{right}}$ is also uniformly distributed on that same set.
Hence $L_{\text{left}} \equiv L_{\text{right}}$ independent of the choice of
$n$.

\subsubsection{2x \% n}

Note that, for $n$ even, $L_{\text{right}}$ is uniformly distributed on:

\[
	2 \cdot Z_n \mod n = \{0, 2, \cdots, n - 2, n, n + 2, \cdots, 2n - 2\} \mod n = \{0, 2, \cdots, n - 2, 0, 2, \cdots, n - 2\}.
\]

For $n$ odd, $L_{\text{right}}$ is uniformly distributed on:

\[
	2 \cdot Z_n \mod n = \{0, 2, \cdots, n - 1, n + 1, \cdots, 2n - 2\} \mod n = \{0, 2, \cdots, n - 1, 1, 3, \cdots, n - 2\}.
\]

Clearly, for $n$ odd, the output follows the same distribution as
$L_{\text{left}}$, so $L_{\text{left}} \equiv L_{\text{right}}$. For $n$ even
however, $L_{\text{right}}$ is uniformly distributed on $\{0, 2, \cdots,
n-2\}$.

A distinguisher $A$ can return $1$ if the given value is even, $0$ if it is
odd. Then, $P[A \diamond L_{\text{left}} \Rightarrow 1] = \frac{1}{2} \neq 1 =
P[A \diamond L_{\text{right}} \Rightarrow 1]$. Hence $L_{\text{left}}$ and
$L_{\text{right}}$ are not exchangable.

\subsubsection{x AND y}

As before, $L_{\text{right}}$ is uniformly distributed on $\{0, 1\}^n$. By the
definition of the bitwise AND operation, for any bit $x_i$ in the output,
$P[x_i = 0] = \frac{3}{4}$, and $P[x_i = 1] = \frac{1}{4}$.

Let $k$ be the number of bits equal to $1$. In the case of $L_{\text{left}}$, k
follows a binomial distribution $Binom(n, \frac{1}{4})$, in the case of
$L_{\text{right}}$ it follows a binomial distribution $Binom(n, \frac{1}{2})$.

This allows solving for the value of $k$ at which the likelihood of the output
$x$ being produced by $L_{\text{left}}$ is equal to it being produced by
$L_{\text{right}}$:

\begin{align*}
	& \binom{n}{k'} \left(\frac{1}{4}\right)^{k'} \left(\frac{3}{4}\right)^{n - k'} = \binom{n}{k'} \left(\frac{1}{2}\right)^n \\
	\Rightarrow & \left(\frac{1}{4}\right)^{k'} \cdot \left(\frac{3}{4}\right)^{n - k'} = \left(\frac{1}{2}\right)^n \\
	\Rightarrow & k' = \frac{n \cdot \log{\frac{3}{2}}}{\log{3}}
\end{align*}

A distinguisher $A$ can then output $1$ if $k > k'$, and $0$ if $k \leq k'$. It
then follows that $P[A \diamond L_{\text{left}} \Rightarrow 1] \leq P[A
\diamond L_{\text{right}} \Rightarrow 1]$, with equality only for special
values of $n$ and $k$ where the equation above yields an integer result.

\section{Security of a modified One-time Pad (OTP)}

We aim to show the one-time secrecy of the given algorithm by showing that
$L_{\text{OTS-left}} \equiv L_{\text{OTS-right}}$, with the two libraries
defined as in the lecture.

For any two messages $m_1, m_2$, the output of the $eavesdrop()$ function of
either library will be a bit-string $c_1 c_2 \cdots c_{n-2} c_{n-1} c_n$, with
$c_{n - 1} = c_n = 0$, and $c_1 c_2 \cdots c_{n-2}$ being uniformly distributed
on $\{0, 1\}^{n-2}$ due to being the result of OTP encryption.

As the ciphertext will thus follow the same distribution in either case, any
distinguishing algorithm $A$ is unable to differentiate between the two
libraries in any way. As such, $P[A \diamond L_{\text{OTS-left}} \Rightarrow 1]
= P[A \diamond L_{\text{OTS-right}} \Rightarrow 1]$, implying that the two
libraries are exchangable, and the given cipher provides one-time secrecy.

\section{Construction of a distinguisher}

Let $m, k, c$ as in the exercise description.

Note that for $m$ even, $k \cdot m \mod 10$ is even no matter the choice of
$k$. For $m$ odd, $k \cdot m \mod 10$ is odd exactly if $k$ is odd, and even
otherwise.
\\

\begin{tabular}{ccc}
	\toprule
	         & $k$ even & $k$ odd \\
	\midrule
	$m$ even & $c$ even & $c$ even \\
	$m$ odd  & $c$ even & $c$ odd \\
	\bottomrule
\end{tabular}
\\

We now define the distinguishing algorithm $A$ as follows:

\begin{lstlisting}[mathescape=true]
	A:
	  m_l <- 2
	  m_r <- 3

	  c = Eavesdrop(m_l, m_r)

	  if c mod 2 == 0
	    # 2/3 likelihood that m_l encrypted
	    return 0
	  else
	    # Guaranteed that m_r encrypted
	    return 1
	  end
\end{lstlisting}

It then follows that $P[A \diamond L_{\text{OTS-left}} \Rightarrow 1] = 0 \neq
\frac{1}{2} = P[A \diamond L_{\text{OTS-right}} \Rightarrow 1]$, hence the two
libraries are not exchangable, and the encryption scheme does not provide
one-time secrecy.

\end{document}

